Turbulence and Particle Acceleration in Giant Radio Halos: the Origin of Seed Electrons

### 2016-04-06 paper.tex
intro modified

There have been two competing models proposed to explain RHs.  The
radio emitting electrons in the ``hadronic model'' are produced in
inelastic (hadronic) CR proton interactions with protons of the
ambient thermal ICM
(\citealp{1980ApJ...239L..93D,1999APh....12..169B,2001ApJ...562..233M,
  2004A&A...413...17P,2008MNRAS.385.1211P,ensslin11}). CR protons and
heavier nuclei may have been accelerated and injected into the ICM by
structure formation shocks, active galactic nuclei and galactic
winds. However, the strong bimodality that separates X-ray luminous
clusters into radio-active and radio-quite clusters (and requires a
fast switch on/off mechanism of the RH emission) and the very extended
RH emission at low frequencies in Coma (352~MHz) represent a major
challenge to this model class \citep{brunetti12,2014MNRAS.438..124Z}.

@@

However, there is a salient piece missing in the turbulent
reacceleration model. It relies heavily on the assumption of an
abundant, volume-filling population of seed suprathermal electrons;
direct Fermi II acceleration from the thermal pool is precluded by
strong Coulomb losses
\citep{2008ApJ...682..175P,2012ApJ...759..113C}. These seeds are
presumed to be either fossil CR electrons (CRes) accelerated by
diffusive shock acceleration (DSA) during structure formation
\citep{1999ApJ...520..529S}, or secondaries injected by hadronic
interaction of CR protons (CRps) with thermal protons
\citep{brunetti11}. While analytic estimates have been made, there has
been no ab initio demonstration that structure formation can lead to
the required abundance of seed electrons with the correct spatial and
spectral characteristics. This is a non-trivial requirement: Coulomb
cooling in dense cluster cores is severe, and DSA fossil electrons may
not survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
\AP{the left panel of} Fig.~\ref{fig:Edens} we show that such a
distribution is not predicted by cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}. \CP{Explain Fig 1: why is
  simulated profile steep? transport, secondaries, \ldots}

@@

Indeed, arriving at a seed population with the required
characteristics is highly constraining, and has the potential to teach
us much about the origin of CRps/CRes in clusters.  \CP{We use our
  simuations to generate distribution function of seed population,
  resolve injection and loss processes for CRs and integrate the
  Fokker-Planck equation along Lagrangian particle trajectories to
  obtain a momentum spectrum of CRs. However, we assume a simplified
  turbulent model. So our approach is orthogonal (and complementary)
  to e.g. Miniati 2015.}  \AP{For both the turbulence and magnetic
  fields we adopt stationary models, parameterized in terms of thermal
  quantities derived from our simulations.}

@@

We incorporate momentum diffusion for electrons and protons from
transit-time-damping (TTD) resonance with compressible
magneto-hydrodynamic (MHD) turbulence, to model Fermi-II
reacceleration \citep{brunetti07,brunetti11}. The TTD resonance
requires the wave frequency $\omega=k_\parallel\upsilon_\parallel$,
where $k_\parallel$ and $\upsilon_\parallel$ are the parallel
(projected along the magnetic field) wavenumber and particle velocity,
respectively. This implies that the particle transit time across the
confining wave region matches the wave period,
$\lambda_{\parallel}/v_{\parallel}=T$. Note, the CRs' gyroradius does
not enter the resonance condition. Hence the CRs that are in resonance
with compressible waves experience Fermi-II acceleration
irrespectively of the length scale of the perturbation.

>>>>

There have been two competing models proposed to explain RHs.  The
radio emitting electrons in the ``hadronic model'' are produced in
inelastic (hadronic) CR proton interactions with protons of the
ambient thermal ICM, which generates pions that eventually decay into
electrons and positrons, depending of the charge of the initial pion
(\citealp{1980ApJ...239L..93D,1999APh....12..169B,2001ApJ...562..233M,
  2004A&A...413...17P,2008MNRAS.385.1211P,ensslin11}). CR protons and
heavier nuclei may have been accelerated and injected into the ICM by
structure formation shocks, active galactic nuclei and galactic
winds. However, the strong bimodality that separates X-ray luminous
clusters into radio-active and radio-quite clusters (and requires a
fast switch on/off mechanism of the RH emission) and the very extended
RH emission at low frequencies in Coma (352~MHz) represent a major
challenge to this model class \citep{brunetti12,2014MNRAS.438..124Z}.

@@

However, there is a salient piece missing in the turbulent
reacceleration model. It relies heavily on the assumption of an
abundant, volume-filling population of seed suprathermal electrons;
direct Fermi II acceleration from the thermal pool is precluded by
strong Coulomb losses
\citep{2008ApJ...682..175P,2012ApJ...759..113C}. These seeds are
presumed to be either fossil CR electrons (CRes) accelerated by
diffusive shock acceleration (DSA) during structure formation
\citep{1999ApJ...520..529S}, or secondaries injected by hadronic
interaction of CR protons (CRps) with thermal protons
\citep{brunetti11}. While analytic estimates have been made, there has
been no ab initio demonstration that structure formation can lead to
the required abundance of seed electrons with the correct spatial and
spectral characteristics. This is a non-trivial requirement: Coulomb
cooling in dense cluster cores is severe, and DSA fossil electrons may
not survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
\AP{the left panel of} Fig.~\ref{fig:Edens} we show that such a
distribution is not predicted by cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}. \CP{If CRps are predominantly
  advected with the cluster plasma, their distribution will be peaked
  towards the cluster center and shows a similar characteristics as
  the thermal plasma. As a consequence, the distribution of secondary
  electrons and the resulting radio synchrotron emission is also
  peaked since the hadronic reaction is a two-body scattering process
  (see Fig.~Fig.~\ref{fig:Edens}). Hence, the simulated emission falls
  short of the observed extended and flat radio profile of the Coma
  cluster.}

@@

Indeed, arriving at a seed population with the required
characteristics is highly constraining, and has the potential to teach
us much about the origin of CRps/CRes in clusters.  \CP{In this work,
  we use our hydrodynamical zoom simulations of galaxy clusters in a
  cosmological setting to follow the distribution functions of seed
  CRp and CRe population and integrate the Fokker-Planck equation of
  CR transport along Lagrangian particle trajectories. We resolve
  injection at structure formation shocks, account for various loss
  processes of CRs, and -- most importantly -- model second-order
  Fermi acceleration by CR interactions with magnetised
  turbulence. However, we assume a simplified and stationary model for
  magnetic fields and turbulence. We do not account for the
  time-varying energy density in compressible waves, which are thought
  to be necessary for the acceleration process
  \citep{brunetti07,brunetti11}, as the cluster merger proceeds.  So
  our approach is orthogonal (and complementary) to e.g.,
  \citealp{2015ApJ...800...60M}. This enables us to vary parameters
  associated with the unknown continuation of magneto-hydrodynamical
  turbulence below the MHD scale and the overall amplitude of
  compressible waves.}

@@

We incorporate momentum diffusion for electrons and protons from
transit-time-damping (TTD) resonance with compressible MHD turbulence,
to model Fermi-II reacceleration \citep{brunetti07,brunetti11}. The
TTD resonance requires the wave frequency
$\omega=k_\parallel\upsilon_\parallel$, where $k_\parallel$ and
$\upsilon_\parallel$ are the parallel (projected along the magnetic
field) wavenumber and particle velocity, respectively. This implies
that the particle transit time across the confining wave region
matches the wave period, $\lambda_{\parallel}/v_{\parallel}=T$. Note,
the CRs' gyroradius does not enter the resonance condition. Hence the
CRs that are in resonance with compressible waves experience Fermi-II
acceleration irrespectively of the length scale of the perturbation.
